---
title: "GTFS Creation for GoBus Service in OHIO"
author: "Orhan Aktas"
date: '2022-11-02'
output:   
    html_document:
      toc: yes
      number_sections: yes
      toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  class.source = "bg-warning text-warning"
)
```

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidytransit)
library(jsonlite)
library(printr)
library(hms)
library(lubridate)
library(readxl)
source("functions.R")
library(rvest)
library(tidygeocoder)
```

# Introduction

In this document, a validated, frequency-based GTFS feed for Athens, Ohio is created from the current informal transit network files. The final output GTFS feed includes the following files that have the required fields.

-   agency.txt

-   stops.txt

-   routes.txt

-   trips.txt

-   stop_times.txt

-   calendar.txt

-   frequencies.txt

-   shapes.txt

# Import data

## Import stops
### Scrape stops from web, create coordinates and save as local file

scrape stops
```{r eval=FALSE}

# stops link
html <- read_html("https://ridegobus.com/stops/")

# get stop names
stop_names <- html %>%
                html_elements(".column-3 a") %>%
                html_text2() %>%
                as_tibble() %>% 
                set_names("stop")

# get stop links
stop_links <- paste0(
                "https://ridegobus.com",
                html %>%
                  html_elements(".column-3 a") %>%
                  html_attr("href") %>%
                  as.character()
              )



```

create function to get stop addresses
```{r eval=FALSE}

get_address <- function(stop_link) {
            address<-read_html(stop_link)%>% 
            html_element(".col-xs-12 :nth-child(1) .col-10 , .col-xs-12 :nth-child(1) strong") %>%     
            html_text2() %>% 
            str_extract_all("(?<=\n).+(?=\n)") %>% 
            pluck(1)
            
            return(address)
}

```

get stop addresses
```{r eval=FALSE}

stop_addresses<-stop_links %>%  
                  map(get_address) %>% 
                  unlist() %>% 
                  as_tibble() %>% 
                  set_names("address") %>% 
                  bind_cols(stop_names) %>% 
                  relocate(stop)
```

save stop address tables
```{r eval=FALSE}
stop_addresses %>% saveRDS("data/ohio/stop_addresses.RDS")
rm(stop_addresses,stop_names,stop_links)
```

add stop coordinates from open street map with geocode function from tidygeocoder package
```{r eval=FALSE}
stop_addresses <- readRDS("data/ohio/stop_addresses.RDS")

stop_addresses %>%
  geocode(address, method = 'osm', lat = latitude , long = longitude)%>%
  saveRDS("data/ohio/lat_longs_missing.RDS")
```

completing missing stops
```{r eval=FALSE}

# read lat_longs table that has missing coordinates
lat_longs_missing <- readRDS("data/ohio/lat_longs_missing.RDS")

# import missing stops that were previously created manually.
missing_stops <- read_excel("data/ohio/missing_stops.xlsx")

# fill the empty missing values from missing stops table
lat_longs <- lat_longs_missing %>%
  left_join(missing_stops, by = "stop") %>%
  mutate(
    latitude = if_else(is.na(latitude), lat, latitude), 
    longitude = if_else(is.na(longitude), lon, longitude)
  ) %>%
  select(-lat, -lon)

# save lat_longs which has no missing coordinates
lat_longs %>% saveRDS("data/ohio/stop_coordinates.RDS")
```


