---
title: "GTFS Creation for Athens"
author: "Orhan Aktas"
date: '2022-05-19'
output:   
    html_document:
      toc: yes
      number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  class.source = "bg-warning text-warning"
)
```

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidytransit)
library(printr)
library(sf)
library(hms)
library(lubridate)
library(units)
library(zip)
library(jsonlite)
library(readxl)
source("functions.R")
```

# Introduction

In this document, a validated, frequency-based GTFS feed for Athens, Ohio is created from the current informal transit network files. The final output GTFS feed includes the following files that have the required fields.

-   agency.txt

-   stops.txt

-   routes.txt

-   trips.txt

-   stop_times.txt

-   calendar.txt

-   frequencies.txt

-   shapes.txt

# Import data

## Sample GTFS data format

```{r import sample_GTFS, message=FALSE, warning=FALSE}
sample_GTFS <- read_gtfs("data/sample-feed.zip")
```

## Current GTFS data format

```{r import current_GTFS, message=FALSE, warning=FALSE}
current_GTFS <- read_gtfs("data/athens.zip")
```

```{r}
GTFS_shapes<-current_GTFS[["shapes"]]
GTFS_stop_times<-current_GTFS[["stop_times"]]
GTFS_stops<-current_GTFS[["stops"]]

GTFS_stop_times %>% 
  left_join(GTFS_stops,by="stop_id") %>% 
  select(trip=trip_id,
         seq=stop_sequence,
         time_a=arrival_time,
         time_d=departure_time,
         stop=stop_id,
         stop_name) %>% View
```

## Transit data from athens.doublemap.com

### import data from webpage

The web information was initially imported from web and then saved to a local directory. Then it is imported from that directory every time the code is executed. The reason for adopting such a process is that the web information is not stable and all of the routes are not published together all the time. So the routes were imported when all the routes were published.

The code below is the initial one for importing from web and saving to the local directory. It is not executed but left in this file for future uses or investigations.

```{r eval=FALSE, include=FALSE}

stops_link<-"https://athens.doublemap.com/map/v2/stops"
trips_link<-"https://athens.doublemap.com/map/v2/trips"
routes_link<-"https://athens.doublemap.com/map/v2/routes"

web_stops <- fromJSON(stops_link, flatten = TRUE)
web_trips<-fromJSON(trips_link, flatten = TRUE)
web_routes<-fromJSON(routes_link, flatten = TRUE)

web_data<-list(web_stops=web_stops,web_trips=web_trips,web_routes=web_routes)
write_rds(web_data,"data/web_data")

rm(stops_link,trips_link,routes_link,web_data,web_stops,web_routes,web_trips)
```

### import web data from local directory

```{r}
web_stops <- read_rds("data/web_data") %>% pluck("web_stops")
web_trips<-read_rds("data/web_data") %>% pluck("web_trips")
web_routes<-read_rds("data/web_data") %>% pluck("web_routes")
```

# create shapes

```{r shapes}
final_shapes<-web_routes %>%
  select(id, path) %>%
  as_tibble() %>%
  unnest(cols = path) %>% 
  rename(shape_pt_lat = path,shape_id=id) %>%
  mutate(shape_pt_lon = lead(shape_pt_lat)) %>%
  filter(shape_pt_lon<0)   %>% 
  group_by(shape_id) %>%
  mutate(shape_pt_sequence = as.integer(row_number()-1)) %>% 
  ungroup()
```

similarity with initial GTFS feed

```{r}
source_gtfs_sum <- GTFS_shapes %>%
  group_by(shape_id) %>%
  summarise(count = n()) %>%
  mutate(shape_id = as.integer(shape_id))

final_shapes %>%
  group_by(shape_id) %>%
  summarise(count = n()) %>%
  left_join(source_gtfs_sum, by = "shape_id")

rm(source_gtfs_sum)
```

It is concluded that although these two files are not same, they are very similar.

# create stops

required fields for stops.txt are the following ones.

-   stop_id

-   stop_name

-   stop_lat

-   stop_lon

Since there are a lot more stops in the web_stops file than the ones in the initial GTFS stop file, the former one is decided to be used.

```{r}
final_stops<-web_stops %>% 
  as_tibble() %>% 
  select(stop_id=id,
         stop_name=name,
         stop_lat=lat,
         stop_lon=lon)
```

# create routes

the fields that are included in the shapes.txt are the following ones.

-   route_id

-   route_type

-   route_long_name (conditionally)

-   route_short_name (conditionally)

-   route_color (optional)

```{r}
final_routes<-web_routes %>% 
  as_tibble() %>% 
  select(route_id=id,
         route_long_name=name,
         route_short_name=short_name,
         route_color =color
         ) %>% 
  mutate(route_type=3L, # for bus mode
         route_long_name=ifelse(route_id==76,"Line 7n",route_long_name),
         route_long_name=ifelse(route_id==78,"Line 7a",route_long_name),
         route_short_name=ifelse(route_id==76,"7n",route_short_name),
         route_short_name=ifelse(route_id==78,"7a",route_short_name)) 
```

# create stop_times

The stop times will be imported from pdf files so the timetables were initially copied to excel files. However these files are not applicaple directly because the stop names are not same with the stop database. Therefore there should be an initial check before using those timetable data.

-   trip_id

-   arrival_time

-   departure_time

-   stop_id

-   stop_sequence

## create input table for timetables

```{r create input table for timetables, eval=FALSE, include=FALSE}
timetable_input_table<-web_routes %>%
  select(route_id=id, stop_id=stops) %>%
  as_tibble() %>%
  unnest(cols = stop_id) %>% 
  group_by(route_id) %>%
  mutate(stop_sequence = as.integer(row_number()-1)) %>% 
  ungroup() %>% 
  left_join(final_stops,by="stop_id") %>% 
  select(-stop_lat,-stop_lon) %>% 
  left_join(final_routes,by="route_id") %>% 
  select(route_id,route_long_name,stop_sequence,stop_id,stop_name) 

openxlsx::write.xlsx(timetable_input_table,file="data/timetable_input_table.xlsx")

rm(timetable_input_table)
```

there are some edits on the files while manipulating the stop_times table

```{r add Clark station}
final_stops<-final_stops %>% add_row(stop_id=182, stop_name="Clarke", stop_lat=39.33556,stop_lon=-82.11262) %>% arrange(desc(stop_id))
```

## import the timetable table

```{r}
timetable_output_table <- read_excel("data/timetable_output_table.xlsx", 
    col_types = c("numeric", "text", "numeric", 
        "numeric", "text", "numeric", "numeric", 
        "skip", "skip"))
```

## create final table

```{r create final_stop_times}
final_stop_times<-timetable_output_table %>%
  mutate(duration_arr = ifelse(stop_sequence != 0, arrival_time - lag(arrival_time), 0)) %>%
  mutate(duration_arr = ifelse(duration_arr < 0, duration_arr + 60, duration_arr)) %>%
  mutate(duration_dep = ifelse(stop_sequence != 0, departure_time - lag(departure_time), 0)) %>%
  mutate(duration_dep = ifelse(duration_dep < 0, duration_dep + 60, duration_dep)) %>%
  group_by(route_id) %>%
  mutate(
    arrival_time_new = round(cumsum(duration_arr), 0),
    departure_time_new = round(cumsum(duration_dep), 0)
  ) %>%
  mutate(
    arrival_time_new = hms::hms(seconds_to_period(arrival_time_new*60)),
    departure_time_new = hms::hms(seconds_to_period(departure_time_new*60))
  ) %>%
  select(
    trip_id = route_id,
    stop_sequence,
    stop_id,
    arrival_time = arrival_time_new,
    departure_time = departure_time_new
  ) %>% 
  ungroup()

rm(timetable_output_table)

final_stop_times %>% head()
```

# create frequencies

Since there are too many records to add to the frequency table and the code will be too dense and harder to read, it is decided to put this code as a function in a separate R script along with similar functions. They are then called at the top of this document.

```{r add_frequency_records}
final_frequencies<-sample_GTFS %>% add_frequency_records()
final_frequencies
```

# create calendar

```{r calendar}
final_calendar<-sample_GTFS %>%
    pluck("calendar") %>%
    slice(0) %>%
  add_row(service_id="MonFri",
          monday=1,
          tuesday=1,
          wednesday=1,
          thursday=1,
          friday=1,
          saturday=0,
          sunday=0,
          start_date=ymd(20220101),
          end_date=ymd(20251231)  
  ) %>% 
    add_row(service_id="MonSat",
          monday=1,
          tuesday=1,
          wednesday=1,
          thursday=1,
          friday=1,
          saturday=1,
          sunday=0,
          start_date=ymd(20220101),
          end_date=ymd(20251231)   
  ) 
final_calendar
  
```

# create agency

```{r}
final_agency<-sample_GTFS %>%
    pluck("agency") %>%
    slice(0) %>% add_row(agency_id="APT",
                         agency_name="Athens Public Transit",
                         agency_url="https://hapcap.org/transportation/athens-public-transit/",
                         agency_timezone="US/Eastern")
final_agency
```

# create trips

```{r}
temp_records<-final_frequencies %>% group_by(trip_id) %>% summarise(count=n()) %>% select(trip_id) %>% mutate(trip_id2=trip_id) %>% separate(trip_id2,sep="_",into = c("route_id","service_id"))

final_trips<-sample_GTFS %>%
    pluck("trips") %>%
    slice(0) %>% 
    select(trip_id,route_id,service_id) %>% bind_rows(temp_records)

rm(temp_records)

final_trips %>% head()
```

```{r}

```
